### 2020-05-31 writeup/transfer.tex
sec 3

We measure model dissimilarity as $\norm{\beta_1 - \beta_2}^2$, which is the distance between source and target in the isotropic model.
We derive a sharp threshold when positive transfer transitions to negative transfer, as model dissimilarity increases.
%Based on Theorem \ref{thm_model_shift}, we derive the transition threshold in the following proposition.
Define the following functions
\begin{align*}
	\Psi(\beta_1, \beta_2) = {\ex{\bignorm{\beta_1 - \beta_2}^2}} / {\sigma^2},  \Phi(\rho_1, \rho_2) = \frac{(\rho_1 + \rho_2 - 1)^2}{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}.
\end{align*}
Let $\alpha_{-} = (1 - \rho_1^{-1/2})^4$ and $\alpha_{+} = (1 + \rho_1^{-1/2})^4$.

\begin{proposition}[Task dissimilarity]\label{prop_dist_transition}
	In the isotropic model, suppose that $\rho_1 > 1$ and $\sigma_1 = \sigma_2 = \sigma$.
	Then we have that
	%Whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the ratio between $\Psi(\beta_1, \beta_2)$ and $\Phi(\rho_1, \rho_2)$:
	\squishlist
		\item If $\Psi(\beta_1, \beta_2) < \alpha_{-} \cdot \Phi(\rho_1, \rho_2)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\Psi(\beta_1, \beta_2) \ge \alpha_{+} \cdot \Phi(\rho_1, \rho_2)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{proposition}

@@

\begin{proposition}[Source data size]\label{prop_data_size}
	In the isotropic model, assume that $\rho_1 > 40$ and $\rho_2 > 500$ are fixed constants, $\Psi(\beta_1, \beta_2) > 2/(\rho_2 - 1)$.
	We have that
	\squishlist
		\item If $\rho_1 < \frac{\rho_2-2}{\alpha_+ \cdot \Phi(\beta_1, \beta_2) (\rho_2 - 1) - 1}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\rho_1 > \frac{\rho_2-2}{\alpha_- \cdot \Phi(\beta_1, \beta_2) (\rho_2 - 3) - 1}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{proposition}
Proposition \ref{prop_data_size} shows that as the source data size increases, we observe a transition from positive to negative transfer.
The proof of Proposition \ref{prop_data_size} is similar to Proposition \ref{prop_dist_transition}, and can be found in Appendix \ref{app_proof_31}.

@@

Given several tasks, let the data efficiency ratio be the largest factor $x$ such that the total number of labeled datapoints needed for solving all tasks can be reduced by $x$ while keeping the same performance compared to STL.
%More precisely, suppose we have $n_i$ datapoints for each task, for $i= 1, 2$.
For $i = 1, 2$, let $\hat{\beta}_i^{\MTL}(x)$ denote the estimator trained using only $\alpha n_i$ datapoints from every task.
The data efficiency ratio is defined as
\[ \argmin_{x\in(0, 1)} ~~
		\te_1(\hat{\beta}_1^{\MTL}(x)) + \te_2(\hat{\beta}_2^{\MTL}(x))
		\le \te_1(\hat{\beta}_1^{\STL}) + \te_2(\hat{\beta}_2^{\STL}). \]
We quantify the data efficiency ratio of the isotropic model as follows.

\begin{proposition}[Labeled data efficiency]\label{prop_data_efficiency}
	In the isotropic model, assume that $\rho_1,\rho_2 > 1$ and $\Psi(\beta_1, \beta_2) < \frac{2}{\rho_1} + \frac{2}{\rho_2}$.
	Then the data efficiency ratio is at least \[ \frac{1}{\rho_1 + \rho_2}\cdot \bigbrace{1 + \frac{1}{\rho_1^{-1} + \rho_2^{-1} - 2\Psi(\beta_1, \beta_2)}}. \]
\end{proposition}
The proof of Proposition \ref{prop_data_efficiency} can be found in Appendix \ref{app_proof_data}.

>>>>

We measure model dissimilarity as $\norm{\beta_1 - \beta_2}^2$, which is the distance between source and target in the isotropic model.
We derive a sharp threshold when positive transfer transitions to negative transfer, as model dissimilarity increases.
%Based on Theorem \ref{thm_model_shift}, we derive the transition threshold in the following proposition.
Define the following functions
\begin{align*}
	\Psi(\beta_1, \beta_2) = {\ex{\bignorm{\beta_1 - \beta_2}^2}} / {\sigma^2},  \Phi(\rho_1, \rho_2) = \frac{(\rho_1 + \rho_2 - 1)^2}{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}.
\end{align*}
Let $\gamma_{-} = (1+o(1))\cdot(1 - \rho_1^{-1/2})^4$ and $\gamma_{+} = (1+o(1))(1 + \rho_1^{-1/2})^4$.
Concretely, if $\rho_1 > 40$, then $\gamma_{-} < 2$ and $\gamma_{+} > 1/2$.

\begin{proposition}[Task dissimilarity]\label{prop_dist_transition}
	In the isotropic model, suppose that $\rho_1 > 1$ and $\sigma_1 = \sigma_2 = \sigma$.
	Then we have that
	%Whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the ratio between $\Psi(\beta_1, \beta_2)$ and $\Phi(\rho_1, \rho_2)$:
	\squishlist
		\item If $\Psi(\beta_1, \beta_2) < \gamma_{-} \cdot \Phi(\rho_1, \rho_2)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\Psi(\beta_1, \beta_2) \ge \gamma_{+} \cdot \Phi(\rho_1, \rho_2)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{proposition}

@@

\begin{proposition}[Source data size]\label{prop_data_size}
	In the isotropic model, assume that $\rho_1 > 40$ and $\rho_2 > 500$ are fixed constants, $\Psi(\beta_1, \beta_2) > 2/(\rho_2 - 1)$.
	We have that
	\squishlist
		\item If $\rho_1 < \frac{\rho_2-2}{\gamma_+ \cdot \Psi(\beta_1, \beta_2) (\rho_2 - 1) - 1}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\rho_1 > \frac{\rho_2-2}{\gamma_- \cdot \Psi(\beta_1, \beta_2) (\rho_2 - 3) - 1}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{proposition}
Proposition \ref{prop_data_size} shows that as the source data size increases, we observe a transition from positive to negative transfer.
The proof of Proposition \ref{prop_data_size} is similar to Proposition \ref{prop_dist_transition}, and can be found in Appendix \ref{app_proof_31}.

@@

Given several tasks, let the data efficiency ratio be the largest factor $x$ such that the total number of labeled datapoints needed for solving all tasks can be reduced by $x$ while keeping the same performance compared to STL.
%More precisely, suppose we have $n_i$ datapoints for each task, for $i= 1, 2$.
For $i = 1, 2$, let $\hat{\beta}_i^{\MTL}(x)$ denote the estimator trained using only $x \cdot n_i$ datapoints from every task.
The data efficiency ratio is defined as
\[ \argmin_{x\in(0, 1)} ~~
		\te_1(\hat{\beta}_1^{\MTL}(x)) + \te_2(\hat{\beta}_2^{\MTL}(x))
		\le \te_1(\hat{\beta}_1^{\STL}) + \te_2(\hat{\beta}_2^{\STL}). \]
We quantify the data efficiency ratio of the isotropic model as follows.

\begin{proposition}[Labeled data efficiency]\label{prop_data_efficiency}
	In the isotropic model, assume that $\rho_1,\rho_2 > 10$ and $\Psi(\beta_1, \beta_2) < \frac{2}{\rho_1} + \frac{2}{\rho_2}$.
	Then the data efficiency ratio is at least \[ \frac{1}{\rho_1 + \rho_2}\cdot \bigbrace{1 + \frac{1}{\rho_1^{-1} + \rho_2^{-1} - 5\Psi(\beta_1, \beta_2)}}. \]
\end{proposition}
The proof of Proposition \ref{prop_data_efficiency} can be found in Appendix \ref{app_proof_data}.
